Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-14 paper/paper.tex
once more to the bridge; back to ten pages again

Use of theorem proving to verify vote counting in elections
is an emerging field. The only use of theorem proving and certifying algorithms~\cite{CertAlg} that we are aware of is our own previous work~\cite{GhaleVSTTE} which just verifies a single instance of STV. While the overall approach is similar, the underlying data structure and the algorithmic content which is being formalised does not allow modular specification, verification and synthesis of verifiers for different types of STV: only a single STV algorithm is
being verified.  Adapting to other schemes
requires a considerable effort of re-encoding (sometimes the whole data structure) and non-trivial expertise with HOL4 and CakeML.

@@

Others employ other theorem proving means such Delphin~\cite{Dolphin} or HOL4. Schurmann and DeYoung~\cite{DeYoung:2012:LLV} have pioneered the use of theorem proving in vote counting. They use Linear Logic~\cite{DBLP:journals/apal/Girard93} to specify the semantics of a specific STV which is similar to Victoria STV except that it transfers votes of an eliminated candidate at once. Although the program obtained from the formalisation performs decently, reading their encoding is challenging due to the complexity of Linear Logic syntax. As the expressive power of Linear Logic is limited compared to higher-order logic, their definition of the count action, for example, uses eight auxiliary assertions which further complicates
the encoding and therefore the adaptability and usability of this approach.  Dawson et al~\cite{DBLP:conf/voteid/DawsonGM15} formalise Hare-Clark STV of Tasmania state in the theorem prover HOL4 and verify properties about the algorithm. They however manually transliterate their  HOL4 implementation into an ML program. There are several issues with such an approach. For example, manual code translation is not a verified process and the semantics of the source (HOL4) and
the target (ML) differ. Hence verification established for the HOL4 implementation does not provably extend to the
transliterated one.
%Ghale et al design a framework for modular formalisation and
%verification of STV algorithms in Coq. The current work discussed
%in this paper is orthogonal to the Coq framework as it deals with
%verification of the computation performed by the Haskell programs
%extracted from Coq using its  automatic program extraction.
%Moreover the framework developed in Coq cannot cover those STV
%algorithms that use random selection of surplus votes for
%transferring them such as the Senate STV of the New South Wales
%state in Australia. However, our framework here can perfectly be
%adjusted to verify executions of programs that compute based on
%such algorithms. Pattinson and Tiwari verify the Schulze Voting
%algorithm in Coq and use the extraction mechanisms of

>>>>

Use of theorem proving to verify vote counting 
is an emerging field. The only use of theorem proving and certifying algorithms~\cite{CertAlg} that we are aware of is our own previous work~\cite{GhaleVSTTE} which just verifies a single instance of STV. While the overall approach is similar, the underlying data structure and the algorithmic content which is being formalised does not allow modular specification, verification and synthesis of verifiers for different types of STV: only a single STV algorithm is
being verified.  Adapting to other schemes
requires a considerable effort of re-encoding (sometimes the whole data structure) and non-trivial expertise with HOL4 and CakeML.

@@

Others employ other theorem proving means such Delphin~\cite{Dolphin} or HOL4. Schurmann and DeYoung~\cite{DeYoung:2012:LLV} have pioneered the use of theorem proving in vote counting. They use Linear Logic~\cite{DBLP:journals/apal/Girard93} to specify the semantics of a specific STV which is similar to Victoria STV except that it transfers votes of an eliminated candidate at once. Although the program obtained from the formalisation performs decently, reading their encoding is challenging due to the complexity of Linear Logic syntax. As the expressive power of Linear Logic is limited compared to higher-order logic, their definition of the count action, for example, uses eight auxiliary assertions which further complicates
the encoding and therefore the adaptability and usability of this approach.  Dawson et al~\cite{DBLP:conf/voteid/DawsonGM15} formalise Hare-Clark STV of Tasmania state in the theorem prover HOL4 and verify properties about the algorithm. They however manually transliterate their  HOL4 implementation into an ML program. There are several issues with this approach. For example, manual code translation is not a verified process. Also the semantics of the source (HOL4) and
the target (ML) differ. Hence verification established for the HOL4 program does not provably extend to the
transliterated one.
%Ghale et al design a framework for modular formalisation and
%verification of STV algorithms in Coq. The current work discussed
%in this paper is orthogonal to the Coq framework as it deals with
%verification of the computation performed by the Haskell programs
%extracted from Coq using its  automatic program extraction.
%Moreover the framework developed in Coq cannot cover those STV
%algorithms that use random selection of surplus votes for
%transferring them such as the Senate STV of the New South Wales
%state in Australia. However, our framework here can perfectly be
%adjusted to verify executions of programs that compute based on
%such algorithms. Pattinson and Tiwari verify the Schulze Voting
%algorithm in Coq and use the extraction mechanisms of
